---
title: "validamiiid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{validamiiid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`{validamiiid}` is a `{shiny}` application that helps experimenting with data that follows the Minimal Information about a Intermicrobial Interaction Data (MIIID) specifications. It uses the JSON schema artefact exported from the MIIID schema developed with the LinkML framework.

Data is written in YAML but validated in JSON which requires a bit of transformation.

The backend of `{validamiiid}` is divided into three functions:

# `yaml_to_json()`

This function converts a YAML string and convert into JSON string.


```{r fct_yaml_to_json, echo=FALSE}
library(yaml)
library(magrittr)
library(jsonlite)
yaml_to_json <- function(yaml_text){
  # Convert a text representation of a YAML to JSON string
  # Trying to limit writing files in the shiny app
  yaml::read_yaml(text = yaml_text) %>% 
    jsonlite::toJSON(., auto_unbox = TRUE) %>% 
    as.character()
}
```


```{r examples_yaml_to_json}
yaml_to_json("")
ex_yaml_pairwise <- paste(c(
  "participants:",
  "  - 'Phocaeicola alice'",
  "  - 'Phocaeicola bob'",
  "tax_id:",
  "  - 353",
  "  - 432"
), collapse = "\n")
yaml_to_json(ex_yaml_pairwise)
```

# `json_to_yaml()`

This function converts (and lint) a JSON string vector (or file) and convert into JSON string vector (or path)

```{r fct_json_to_yaml, echo=FALSE}
json_to_yaml <- function(json_text, ...){
  # Convert a text representation of a JSON to a YAML string
  jsonlite:::fromJSON(json_text) %>% 
    yaml::as.yaml(., indent.mapping.sequence = TRUE)
}
```

```{r examples_json_to_yaml}
ex_json_pairwise <- '{
  "participants": ["Phocaeicola alice","Phocaeicola bob"],
  "tax_id":[353,432]
}'
json_to_yaml(ex_json_pairwise)
```


# validate MIIID data

This process relies on `{jsonvalidate}` to validate the converted data against the [MIIID schema](https://fair-mi.github.io/miiid-schema).

## Schema generation

A specific artifact of JSON has to be generated from the schema in LinkML:

```bash
gen-json-schema -t IntermicrobialInteraction src/miiid_schema/schema/miiid_schema.yaml > miiid_jsonschema.json
# TODO: look into sub-schema validation using 
# `jsonvalidate::json_validator` option reference = "#/$defs/IntermicrobialInteraction"
# and why it throws a Error: TypeError: validator is not a function
```

## Build validator once

```{r fct_validate_miiid_data}
library(jsonvalidate)
miiid_validator <- jsonvalidate::json_validator(
  system.file("extdata", "miiid_jsonschema.json", package = "validamiiid"),
  engine = "ajv"
)
```


## Validate

```{r examples_validate_miiid_data}
# Load examples
data("valid_example_miiid", package = "validamiiid")
data("invalid_example_miiid", package = "validamiiid")

# Conversion to JSON
valid_json <- yaml_to_json(valid_example_miiid)
invalid_json <- yaml_to_json(invalid_example_miiid)

miiid_validator(valid_json)
miiid_validator(invalid_json)
```

## Parse validator

```{r fct_parse_validator}
parse_validator <- function(validator_result) {
  if (is.null(attr(validator_result, "errors"))) {
    stop("This validator result is not valid. Please run the validator with verbose = TRUE.")
  }
  errors_df <- attr(validator_result, "errors")
  return(errors_df[, c("instancePath", "message")])
}
```

```{r examples_parse_validator}
res <- miiid_validator(invalid_json, verbose = TRUE)
parse_validator(res)
```


# functions to add

## `yaml_to_gh_issue()`

This is to format/encode a URL with a template for an issue that would carry over a YAML file to Github for further examination. This function will be run only when validation was carried out.

```{r fct_yaml_to_gh_issue, echo=FALSE}
yaml_to_gh_issue <- function(
    yaml_text,
    repo_url = "https://github.com/fair-mi/miiid-schema",
    assignees = "cpauvert",
    label = "miiid-sandbox",
    title = "Description of the interaction <INTERACTION_ID>",
    pre_yaml_template = base::paste(
      "<!--   Thank you for contributing!   -->",
      "<!--   Please edit the information below to suit your description   -->",
      "The following `YAML` file describes the interaction: <INTERACTION_ID>.",
      "", "```yaml",
      sep = "\n"
    ),
    post_yaml_template = base::paste(
      "```", "",
      "<!--   Your contribution is labelled as sandbox by default   -->",
      "<!--   If you want more than trying out:   -->",
      "<!--      you can label this issue miiid-contribution   -->",
      "<!--      instead of miiid-sandbox for further consideration   -->", "",
      "<!--   Feel free to add other relevant information,   -->",
      "<!--    and check the Preview tab ;)   -->",
      sep = "\n"
    )) {
  # Assemble the template together with the YAML text
  tmplt <- base::paste0(
    pre_yaml_template, "\n",
    yaml_text,
    post_yaml_template
  )
  # Extract the interaction identifier
  interaction_id <- base::gsub("^id: (.+)\\nparticipants.*", "\\1", yaml_text)
  interaction_id <- base::paste0("`", interaction_id, "`")
  title <- base::gsub("<INTERACTION_ID>", interaction_id, title)
  tmplt <- base::gsub("<INTERACTION_ID>", interaction_id, tmplt)

  # Concatenate the full URL
  issue_url <- base::paste0(
    repo_url, "/issues/new?",
    "assignees=", assignees,
    "&labels=", label,
    "&title=", title,
    "&body=", tmplt
  )
  return(utils::URLencode(issue_url))
}
```


```{r ex_yaml_to_gh_issue}
yaml_to_gh_issue( json_to_yaml(ex_json_pairwise) )
```

## `get_schema_version()`

Extract the release version of the MIIID schema used for validation.

```{r fct_get_schema_version, echo=FALSE}
get_schema_version <- function() {
  jsonlite::read_json(
    system.file(
      "extdata", "miiid_jsonschema.json",
      package = "validamiiid"
    )
  )[["version"]]
}
```

```{r ex_get_schema_version}
get_schema_version()
```
