---
title: "valida-miiid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{valida-miiid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`{valida-miiid}` is a `{shiny}` application that helps experimenting with data that follows the Minimal Information about a Intermicrobial Interaction Data (MIIID) specifications. It uses the JSON schema artefact exported from the MIIID schema developed with the LinkML framework.

Data is written in YAML but validated in JSON which requires a bit of transformation.

The backend of `{valida-miiid}` is divided into three functions:

# `yaml_to_json()`

This function converts a YAML string and convert into JSON string.


```{r fct_yaml_to_json, echo=FALSE}
library(yaml)
library(magrittr)
library(jsonlite)
yaml_to_json <- function(yaml_text){
  # Convert a text representation of a YAML to JSON string
  # Trying to limit writing files in the shiny app
  yaml::read_yaml(text = yaml_text) %>% 
    jsonlite::toJSON(., auto_unbox = TRUE)
}
```


```{r examples_yaml_to_json}
yaml_to_json("")
ex_yaml_pairwise <- paste(c(
  "participants:",
  "  - 'Phocaeicola alice'",
  "  - 'Phocaeicola bob'",
  "tax_id:",
  "  - 353",
  "  - 432"
), collapse = "\n")
yaml_to_json(ex_yaml_pairwise)
```

# `json_to_yaml()`

This function converts (and lint) a JSON string vector (or file) and convert into JSON string vector (or path)

```{r}
json_to_yaml <- function(json_text, ...){
  # Convert a text representation of a JSON to a YAML string
  jsonlite:::fromJSON(json_text) %>% 
    yaml::as.yaml(., indent.mapping.sequence = TRUE)
}
```

```{r examples_json_to_yaml}
ex_json_pairwise <- '{
  "participants": ["Phocaeicola alice","Phocaeicola bob"],
  "tax_id":[353,432]
}'
json_to_yaml(ex_json_pairwise)
```


# `validate_miiid_data()`

This function relies on `{jsonvalidate}` to validate the converted data against the [MIIID schema](https://fair-mi.github.io/miiid-schema).

## Schema generation

A specific artifact of JSON has to be generated from the schema in LinkML:

```bash
gen-json-schema -t IntermicrobialInteraction src/miiid_schema/schema/miiid_schema.yaml > miiid_jsonschema.json
# TODO: look into sub-schema validation using 
# `jsonvalidate::json_validator` option reference = "#/$defs/IntermicrobialInteraction"
# and why it throws a Error: TypeError: validator is not a function
```

## Validation

```{r fct_validate_miiid_data}
library(jsonvalidate)
validate_miiid_data <- function(json_text, miiid_schema){
  json_validate(json_text, miiid_schema, verbose = T, engine = "ajv")
  # TODO: capture the tabular output when invalid file is used
}
```


```{r examples_validate_miiid_data}
validate_miiid_data(
  here::here("data/valid.json"),
  here::here("data/miiid_jsonschema.json")
)
validate_miiid_data(
  here::here("data/invalid.json"),
  here::here("data/miiid_jsonschema.json")
)
```


# utils to add

* TODO: add utils function get_schema_version
* ~~TODO: add utils function insert_miiid_type~~
* TODO: add utils function format/encode url for github issue
